---
title: "STATS 415 Final Project"
author: "Margot Douillet, Richard Einhorn, Nathan Nguyen, Edmund Tian"
date: "12/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
set.seed(42)
```

## 2.3 Linear Regression
```{r}
final_project <- read_csv("data/final_project-1.csv")
# Removing the index column
final_project <- final_project %>% select(Asset_1:Asset_3)
Asset_1 <- final_project %>% select(Asset_1)
df <- read.csv("output/bret.csv")
Asset_1_lead <- lead(Asset_1, n=10, default=tail(Asset_1, 1))
Asset_1_HRet_10 <- (Asset_1_lead - Asset_1) / Asset_1
colnames(Asset_1_HRet_10) <- c("Asset_1_HRet_10")
df <- cbind(df, Asset_1_HRet_10)

train_size <- floor(nrow(df) * 0.7)
test_size <- nrow(df) - train_size

train_set <- head(df, train_size)
test_set <- tail(df, test_size)

lr_modl <- lm(Asset_1_HRet_10 ~ ., data=train_set)
summary(lr_modl)
```
It seems that the 3, 10, and 30 minutes backward returns of Asset 2 and the 3 minutes backward return of Asset 3 are important in predicting the forward return of Asset 1.
```{r}
train_pred <- predict.lm(lr_modl, train_set)
test_pred <- predict.lm(lr_modl, test_set)

# In-sample correlation
cor(as.matrix(cbind(train_pred, train_set$Asset_1_HRet_10)))
sprintf("The in-sample correlation is %s", 
        cor(as.matrix(cbind(train_pred, train_set$Asset_1_HRet_10)))[[2,1]])
# Out-sample correlation
cor(as.matrix(cbind(test_pred, test_set$Asset_1_HRet_10)))
sprintf("The out-sample correlation is %s",
        cor(as.matrix(cbind(test_pred, test_set$Asset_1_HRet_10)))[[2,1]])

train_pred <- data.frame(train_pred)
test_pred <- data.frame(test_pred)
colnames(train_pred) <- c("Asset_1_HRet_10_pred")
colnames(test_pred) <- c("Asset_1_HRet_10_pred")
Asset_1_HRet_10_lead <- rbind(train_pred, test_pred)
df <- cbind(df, Asset_1_HRet_10_lead)

# 3 Weeks Rolling correlation
for (i in 1:nrow(df)) {
  start = max(i - 30240, 1)
  df$Rho[i] = cor(df$Asset_1_HRet_10[start:i], df$Asset_1_HRet_10_pred[start:i])
}
plot(df$Rho, xlab="Time", ylab="Correlation", 
     main="3 Weeks Rolling correlation between true and pred")
```
This correlation structure is unstable near the beginning of the year but is relatively stationary for the year. There are some fluctuations but the correlation seem to stay between 0 and 0.1.




## 2.4 KNN Regression

Helper Functions
```{r}
# h-min backward return
h_min_br <- function(df, h){
  data <- df
  a <- h + 2
  c <- h + 1
  data[a:nrow(data),] <- data[2:(nrow(data)-h),]
  data[1:c,] <- data[2,]
  data <- (df - data)/(data)
}

# h-min forward return
h_min_f <- function(df, h){
  data <- df
  a <- (nrow(df) - h)
  b <- h + 1
  c <- a + 1
  d <- nrow(df)
  data[c:d,] <- data[d,]
  data[1:a,] <- data[b:d,]
  data <- (data - df)/df
}
```

Creating Predictors, Testing and Training data
```{r}
# Response vector
rf_t10 <- h_min_f(final_project, 10)[,1]
# Predictors
# Using the h-min backward return from part 2.1
rb_t3 <- h_min_br(final_project, 3)[,1]
rb_t10 <- h_min_br(final_project, 10)[,1]
rb_t30 <- h_min_br(final_project, 30)[,1]
rb_df <- data.frame(rf_t10, rb_t3, rb_t10, rb_t30)
# Splitting the dataframe into training and testing
Train_rb <- rb_df[1:366912,]
Test_rb <- rb_df[366913:524160,]
```

KNN Regression
```{r}
library(FNN)
# The different K-values to be used
K <- c(5,25,125,625,1000)
# Finding the test MSE for each K
testMSE <- c()
for(i in 1:length(K)){
  est <- knn.reg(train = Train_rb[,2:4], test = Test_rb[2:4], y = Train_rb$rf_t10, k = K[i])
  testMSE[i] <- mean((est$pred - Test_rb$rf_t10)^2)
}

# Finding the training MSE for each K
trainMSE <- c()
for(i in 1:length(K)){
  est <- knn.reg(train = Train_rb[,2:4], test = Train_rb[2:4], y = Train_rb$rf_t10, k = K[i])
  trainMSE[i] <- mean((est$pred - Train_rb$rf_t10)^2)
}

# Plotting the test MSEs
plot(K, testMSE, xlab = "K", ylab = "Testing MSE", main = "Testing MSE as a function of K", type = "l", col = "blue4")
# Plotting the training MSEs
plot(K, trainMSE, xlab = "K", ylab = "Training MSE", main = "Training MSE as a function of K", type = "l", col = "red4")

# Finding the K value that yields the lowest training MSE
K[which.min(trainMSE)]
# Finding the K value that yields the lowest test MSE
K[which.min(testMSE)]
```

Finding Training and Test estimates
```{r}
# Performing KNN regression with K = 1000
Train_est <- knn.reg(train = Train_rb[,2:4], test = Train_rb[2:4], y = Train_rb$rf_t10, k = 1000)$pred

Test_est <- knn.reg(train = Train_rb[,2:4], test = Test_rb[2:4], y = Train_rb$rf_t10, k = 1000)$pred

# Finding the in-sample and out-of-sample correlations
cor(Test_est, Test_rb$rf_t10)
cor(Train_est, Train_rb$rf_t10)
```
The correlation between the actual training response and the predicted training response is 0.083, while the correlation between the predicted test response and the actual test response is 0.029. 